\documentclass[../INDE250HW.tex]{subfiles} 
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}

\title{UW IND E 250 Notes}
\author{Anthony Le}

\begin{document}

\pagestyle{fancy}
\fancyhead{}
\fancyhead[R]{UW IND E 250}
\fancyhead[L]{Anthony Le}

\begin{center}
    \LARGE{\textbf{IND E 250 HW 4}}
\end{center}
\begin{center}
    Homework 5 Problems: \\
    7.8, 7.10 (part a and b only), 7.19, 7.23, 7.30, 7.36 (bonus)
\end{center}

\section*{Problem 7.8}
\begin{exmp}
    Consider 4 investments with the sequences of cash flows given in Table P7.8.
    \begin{enumerate}
        \item Identify all the simple investment.
        \item Identify all the nonsimple investments.
        \item Compute i* for each investment.
        \item Which project has no rate of return?
    \end{enumerate}
\end{exmp}



\section*{Problem 7.10}
\begin{exmp}
    Consider the investment projects given in Table P7.10.
    \begin{enumerate}
        \item Classify each project as either simple or non-simple.
        \item Use the quadratic equation to compute i* for project A.
    \end{enumerate}
\end{exmp}


\section*{Problem 7.19}
\begin{exmp}
    Consider the investment projects given in Table P7.19.
    
    Assume that MARR = 12\% in the following questions:
    \begin{enumerate}
        \item Identify the i*(s) for each investment. If the project has more than 1 i*, identify all of them.
        \item Which project(s) is (are) a mixed investment?
    \end{enumerate}
\end{exmp}


\section*{Problem 7.23}
\begin{exmp}
    Consider the investment projects given in Table P7.19.

    Assume MARR = 12\% in the following questions:
    \begin{enumerate}
        \item Compute i* for projects A and B. If the project has more than i*, identify all of them. 
        \item Classify each project as either a pure or mixed investment.
        \item Compute the IRR for each investment.
        \item Determine the acceptability of each project.
    \end{enumerate} 
\end{exmp}


\section*{Problem 7.30}
\begin{exmp}
    Consider the investment projects given in Table P7.30.
    If the project's IRR is 10\%:
    \begin{enumerate}
        \item Find the value of X.
        \item Is this project acceptable at MARR = 8\%?
    \end{enumerate}
\end{exmp}



\end{document}