\documentclass{report} % //TODO find why latex gives errors for report and using textsc!
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[margin=1in]{geometry}

\title{UW CEE 442 Notes}
\author{Anthony Le}

\newtheorem{exmp}{Example}
\newtheorem{exrc}{Excersize}
\newtheorem{proof}{Proof}
\newtheorem{defn}{Definition}


\begin{document}

\pagestyle{fancy}
\fancyhead{}
\fancyhead[R]{UW CEE 442}
\fancyhead[L]{Anthony Le}

\begin{center}
    \LARGE{\textbf{Lecture 2 Notes}}
\end{center}


Today, you'll work in your groups:
\begin{enumerate}
    \item Fill out a canvas quiz
\end{enumerate}

\section*{Canvas Quiz Responses}
\begin{enumerate}
    \item You need to design the slab for the forces from your demand analysis. WHere 
\end{enumerate}
\begin{enumerate}
    \item You need to design the slab for the forces from your demand analysis.Where would you go for information on the design and detailing of concrete slabs? \\
    \textbf{Find this using ACI 318 or in Canvas Files - 06 Gravity Design - Concrete Gravity Design}

    \item You need to design your composite deck for your building which uses a steel gravity frame and a special concentrically braced frame. Where would you find the documents needed to help you? \\
    \textbf{Find this using the AISC Steel Manual or in Canvas Files - 06 Gravity Design - Steel Gravity Design (Look for the Verco Floor Deck Catalog)}

    \item I do not understand how to format my design report. Where should I look? \\
    \textbf{Go to Canvas Files - 10 Writing and Presentation - Writing - Technical Writing Presentation Guidelines}

    \item My group has been assigned the design of a special concentrically braced frame. What documents do you need?\\
    \textbf{Find this using the AISC Steel Manual or in Canvas Files - 07 Lateral Design - Steel/Concrete Design}
    
    \item What is included in dead load? (What document tells you this and where is on Canvas?) \\
    \textbf{Reference the ASCE Manual - reference the architectural plans to find the purpose of the room and the according dead loads according to the room use in ASCE}

    \item Yikes! I need to make a presentation to the class. Can you find supporting documentation on creating and giving presentations? (again, find it on Canvas) \\
    \textbf{Go to Canvas Files - 10 Writing and Presentation - Presentations}

    \item Where is ACI 318-19? Where is AISC 360? Where is ASCE 7? \\
    \textbf{Go to Canvas Files - 09 Code Resources}

\end{enumerate}

\section*{Lecture 4 Notes}
Will be teaching us mostly about BlueBeam and how to use


\end{document}

